### 2017-01-22 intro.tex
changes almost done

The asymptotic convergence of 1SPSA algorithm requires one extra assumption other than the conditions specified in the FDSA scheme above. The extra condition for convergence of $\theta_n$ to $\theta^*$  is as follows:
\begin{pvn}
\item $\{\Delta_{ni}, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., independent of $\F_n$ and for some $\alpha_0 > 0$ and $\forall n$, $\E (\Delta_{ni})^{-2} \le \alpha_0$, for $i = 1,2,\ldots,N$.
\end{pvn}
Since 1SPSA randomly perturbs the parameter vector $\theta$, the number of function measurements required is two, irrespective of the dimension of $\theta$, see that the numerator of \eqref{eq:spsa} is the same in all $N$ components of $\widehat{\nabla} f(\theta_{n})$. The efficiency of 1SPSA depends on the shape of the objective function $f$, $\{a_n\}$, $\{\delta_n\}$ and the distribution of perturbation vectors, i.e., $\Delta_n$ $\forall n$. In \cite{spall}, in addition to establishing formal convergence of $\theta_n \to \theta^*$, it was shown that probability distribution of appropriately scaled $\theta_n$ is approximately normal  (with some mean and covariance matrix) for larger $n$. The general metric which is used for evaluating these class of algorithms is asymptotic mean squared error (AMSE) defined as $\E\l \theta_{n} - \theta^* \r^2$.  The relative efficiency of 1SPSA when compared to FDSA can be shown when we compare the similar asymptotic normality result of FDSA. In \cite{spall}, it was shown that under reasonably general conditions, and when $\{a_n\}$, $\{\delta_n\}$ are chosen according to the guidelines in \cite{spall1998implementation} then 1SPSA and FDSA gives same AMSE for given number of iterations. Instead of looking the 
\begin{align}\label{eq:amse}
\frac{\text{AMSE of 1SPSA}}{\text{AMSE of FDSA}}, 
\end{align}
an equivalent way of looking the relative AMSE is to compare the number of function measurements required by FDSA to achieve same AMSE as 1SPSA which can be computed by equating the \eqref{eq:amse} to one.
\begin{align}\label{eq:amsenom}
 \frac{\text{No.of function measurements for 1SPSA }}{\text{No.of function measurements for FDSA}}=  \frac{1}{N}.
\end{align}

>>>>

The asymptotic convergence of 1SPSA algorithm requires one extra assumption other than the conditions specified in the FDSA scheme above. The extra condition for convergence of $\theta_n$ to $\theta^*$  is as follows:
\begin{pvn}
\item $\{\Delta_{ni}, i=1,\ldots,N, n=1,2,\ldots\}$ are i.i.d., independent of $\F_n$ and for some $\alpha_0 > 0$ and $\forall n$, $\E (\Delta_{ni})^{-2} \le \alpha_0$, for $i = 1,2,\ldots,N$.
\end{pvn}
Since 1SPSA randomly perturbs the parameter vector $\theta$, the number of function measurements required is two, irrespective of the dimension of $\theta$, see that the numerator of \eqref{eq:spsa} is the same in all $N$ components of $\widehat{\nabla} f(\theta_{n})$. The efficiency of 1SPSA depends on the shape of the objective function $f$, $\{a_n\}$, $\{\delta_n\}$ and the distribution of perturbation vectors, i.e., $\Delta_n$ $\forall n$. In \cite{spall}, in addition to establishing formal convergence of $\theta_n \to \theta^*$, it was shown that probability distribution of appropriately scaled $\theta_n$ is approximately normal  (with some mean and covariance matrix) for larger $n$. The general metric which is used for evaluating these class of algorithms is asymptotic mean squared error (AMSE), defined as follows  $\E\l \theta_{n} - \theta^* \r^2$.  The relative efficiency of 1SPSA when compared to FDSA can be shown when we compare the similar asymptotic normality result of FDSA. That is
\begin{align}\label{eq:amse}
\frac{\text{AMSE of 1SPSA}}{\text{AMSE of FDSA}},
\end{align}
In \cite{spall}, it was shown that under reasonably general conditions, and when $\{a_n\}$, $\{\delta_n\}$ are chosen according to the guidelines in \cite{spall1998implementation} then 1SPSA and FDSA gives same AMSE when both the algorithms are ran for the same number of iterations. 
Instead of looking \eqref{eq:amse}, an equivalent way of looking the relative AMSE is to compare the number of function measurements required by FDSA to achieve same AMSE as 1SPSA. One can compute it  by equating the \eqref{eq:amse} to one. Then it will result the following:
\begin{align}\label{eq:amsenom}
 \frac{\text{No.of function measurements for 1SPSA }}{\text{No.of function measurements for FDSA}}=  \frac{1}{N}.
\end{align}

### 2017-01-22 second-order-spsa-3.tex
changes almost done

We next present a convergence rate result for the special case of a quadratic objective function under the following additional assumptions:
\begin{pvn}
\item  $f$ is quadratic and $\nabla^2 f(x^*) > 0$. 
\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and \\$\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
\end{pvn}
\begin{theorem}(\textbf{2SPSA-3-IH Quadratic case - Convergence rate})
\label{thm:spsaquad-bound}
Assume (A23), (A29), (A30) and (A31) and also that the setting is noise-free. 
Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(x^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
\begin{align}
\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}}).
\label{eq:spsaquad-bigo}
\end{align}
\end{theorem}
\begin{proof}
Since the setting is noise-free with a quadratic objective, we can rewrite \eqref{eq:spsahnhat} as follows:
\begin{align}
 \widehat H_n &=    \nabla^2 f(\theta_n) +\Psi_{n}(\nabla^2 f(\theta_n))\nonumber\\
 &=    \nabla^2 f(\theta^*) +\Psi_{n}(\nabla^2 f(\theta_n))\nonumber\\
 &=    H^* +\Psi_{n}(H^* ),\label{eq:spsahnhat-ext}
\end{align}
where $\Psi_n(H)$ is defined in \eqref{eq:psinhatspsa3}.
The proof involves the following steps:
\begin{description}
  \item[Step 1:] Here we prove the MSE convergence of $\overline H_k$, i.e., $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s. as $n\rightarrow \infty$.
  \item[Step 2:] We unroll the recursion \eqref{eq:2spsa-3-ih} and then derive a convenient representation for the  $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
  \item[Step 3:] We derive the main result in \eqref{eq:spsaquad-bigo} using a proof by contradiction. 
\end{description}
\noindent\textbf{Step 1: MSE convergence of $\overline H_n$} \\
This part in exactly the same manner as part (i) in Theorem 3 of \cite{spall-jacobian}.\\
By rewriting the \eqref{eq:2spsa-3-ih} in the following form
\begin{align}\label{eq:hhatexpanded}
\overline H_{n} &= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - \widehat H_n + \widehat \Psi_n)\nonumber\\
&= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - H^* + \widehat \Psi_n - \Psi_n (H^*))
\end{align}
The above equation can be thought as a stochastic approximation analogue of \eqref{eq:2spsa-3-ih}. The above stochastic approximation algorithm aimed at finding the roots of the equation $H-H^*=0$. In \eqref{eq:hhatexpanded}, $\E (\widehat \Psi_n - \Psi_n (H^*) | \F_n) = 0 \,\,\ a.s$.  Given a Lyapunov function, by showing the conditions required to show   the mean square convergence of the \eqref{eq:hhatexpanded} from the \cite{nevel1973stochastic}, pp. 92-94. From part (i) in Theorem 3 of \cite{spall-jacobian}, one can check that our algorithm indeed satisfy those conditions,  which in turn implies the $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s. as $n\rightarrow \infty$.

@@

Now consider the second case where $lim sup_{n \to \infty} \frac{d(n)}{n^{2r}} = \infty$. There exists a subsequence $\{i_1,\ldots,i_n\}$ such that $ \frac{d(i_n)}{i_n^{2r}} \ge \frac{d(i)}{i^{2r}}$ as $i \le i_n$. Then for some $C >0$ the RHS of \eqref{eq:thmwts} satisfies the following 
\begin{align}
\text{RHS}(i_n)  \text{of}  \eqref{eq:thmwts} &= O(1) e^{-2 \Gamma_{1i_n}} +  O(1) e^{-2 b_0 i_n^{1-r}/(1-r)} \left[\sum_{i=1}^{i_n} e^{2 b_0 i^{1-r}/(1-r)} \frac{b_0^2}{i^{2r}} \E (\Lambda_{n} \otimes \Lambda_{n}) \right]\nonumber \\
&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + O(1) \sum_{i =1}^{i_n} \frac{d(i)}{i^{2r}}\right]\nonumber \\
&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + \frac{d(i_n)}{{i_n}^{2r}}\right].
\label{eq:spsaquad-bigocont5}
\end{align}
From the equation \eqref{eq:spsaquad-bigocont}, the  LHS of \eqref{eq:thmwts} is $d(i_n) e^{-2b_0 i_n^{1-r} / {1-r}}$. By substituting \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont5}, in $\frac{RHS(i_n)}{LHS(i_n)}$ then $\frac{RHS(i_n)}{LHS(i_n)} \to 0$ as $n \to \infty$. It is contradiction since $LHS(n) = RHS(n)$ for all $n$. Hence the claim $\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}})$ follows.
\end{proof}

>>>>

We next present a convergence rate result for the special case of a quadratic objective function under the following additional assumptions:
\begin{pvn}
\item  $f$ is quadratic and $\nabla^2 f(x^*) > 0$. 
\item The operator $\Upsilon$ is chosen such that $\E(\parallel \Upsilon(\overline H_n) - \overline H_n\parallel^2) = o(e^{-2wn^{1-r}/(1-r)})$ and \\$\parallel \Upsilon(H) - H \parallel^2 / (1+\parallel H \parallel^2)$ is uniformly bounded.
\end{pvn}
\begin{theorem}(\textbf{2SPSA-3-IH Quadratic case - Convergence rate})
\label{thm:spsaquad-bound}
Assume (A23), (A29), (A30) and (A31) and also that the setting is noise-free. 
Let $b_n = b_0/n^r$, $n=1,2,\ldots,k$, where $1/2 < r < 1$ and $0 < b_0 \leq 1$. For notational simplicity, let $H^*=\nabla^2 f(x^*)$. Letting $\Lambda_k = \overline H_k - H^*$, we have 
\begin{align}
\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}}).
\label{eq:spsaquad-bigo}
\end{align}
\end{theorem}
\begin{proof}
Since the setting is noise-free with a quadratic objective, we can rewrite \eqref{eq:spsahnhat} as follows:
\begin{align}
 \widehat H_n &=    \nabla^2 f(\theta_n) +\Psi_{n}(\nabla^2 f(\theta_n))\nonumber\\
 &=    \nabla^2 f(\theta^*) +\Psi_{n}(\nabla^2 f(\theta^*))\nonumber\\
 &=    H^* +\Psi_{n}(H^* ),\label{eq:spsahnhat-ext}
\end{align}
where $\Psi_n(H)$ is defined in \eqref{eq:psinhatspsa3}.
The proof involves the following steps:
\begin{description}
  \item[Step 1:] Here we prove the MSE convergence of $\overline H_k$, i.e., $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0$ a.s. as $n\rightarrow \infty$.
  \item[Step 2:] We unroll the recursion \eqref{eq:2spsa-3-ih} and then derive a convenient representation for the  $\text{trace}[\E (\Lambda_n \tr \Lambda_n)]$.
  \item[Step 3:] We derive the main result in \eqref{eq:spsaquad-bigo} using a proof by contradiction. 
\end{description}
\noindent\textbf{Step 1: MSE convergence of $\overline H_n$} \\
This part in exactly the same manner as part (i) in Theorem 3 of \cite{spall-jacobian}.\\
By rewriting the \eqref{eq:2spsa-3-ih} in the following form
\begin{align}\label{eq:hhatexpanded}
\overline H_{n} &= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - \widehat H_n + \widehat \Psi_n)\nonumber\\
&= \overline H_{n-1} - b_{n} ( \overline H_{n-1} - H^* + \widehat \Psi_n - \Psi_n (H^*)).
\end{align}
The above equation can be thought as a stochastic approximation analogue of \eqref{eq:2spsa-3-ih}. The above stochastic approximation algorithm aimed at finding the roots of the equation $H-H^*=0$. In \eqref{eq:hhatexpanded}, $\E (\widehat \Psi_n - \Psi_n (H^*) | \F_n) = 0 \,\,\ a.s$.  Given a Lyapunov function, one can  the conditions required to show  the mean square convergence of the \eqref{eq:hhatexpanded} from the \cite{nevel1973stochastic}, pp.92-94. It is easy to see that $V(H) = \frac{1}{2} (H-H^*)\tr(H-H^*)$ indeed serves as Lyapunov function. From part (i) in Theorem 3 of \cite{spall-jacobian}, one can check that our algorithm indeed satisfy those conditions,  which in turn implies the $\E [\Lambda_n\tr \Lambda_n] \rightarrow 0 \,\,\,a.s.$ as $n\rightarrow \infty$.

@@

Now consider the second case where $lim sup_{n \to \infty} \frac{d(n)}{n^{2r}} = \infty$. There exists a subsequence $\{i_1,\ldots,i_n\}$ such that $ \frac{d(i_n)}{i_n^{2r}} \ge \frac{d(i)}{i^{2r}}$ as $i \le i_n$. Then for some $C >0$ the RHS of \eqref{eq:thmwts} satisfies the following 
\begin{align}
\text{RHS}(i_n)\,\,\,  \text{of}  \,\,\,\eqref{eq:thmwts} &= O(1) e^{-2 \Gamma_{1i_n}} +  O(1) e^{-2 b_0 i_n^{1-r}/(1-r)} \left[\sum_{i=1}^{i_n} e^{2 b_0 i^{1-r}/(1-r)} \frac{b_0^2}{i^{2r}} \E (\Lambda_{n} \otimes \Lambda_{n}) \right]\nonumber \\
&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + O(1) \sum_{i =1}^{i_n} \frac{d(i)}{i^{2r}}\right]\nonumber \\
&\le C e^{-2b_0n^{1-r} / {1-r}} \left[ 1 + \frac{d(i_n)}{{i_n}^{2r}}\right].
\label{eq:spsaquad-bigocont5}
\end{align}
Where $\text{RHS}(i_n)$ denotes, RHS after the algorithm is iterated for $i_n$ times. From the equation \eqref{eq:spsaquad-bigocont}, the  LHS of \eqref{eq:thmwts} after $i_n$ iterations is $d(i_n) e^{-2b_0 i_n^{1-r} / {1-r}}$. By substituting \eqref{eq:spsaquad-bigocont} and \eqref{eq:spsaquad-bigocont5}, in $\frac{RHS(i_n)}{LHS(i_n)}$ then $\frac{RHS(i_n)}{LHS(i_n)} \to 0$ as $n \to \infty$. It is contradiction since $LHS(n) = RHS(n)$ for all $n$. Hence the claim $\text{trace}[\E (\Lambda_n \tr \Lambda_n)] = O(e^{-2b_0n^{1-r} / {1-r}})$ follows.
\end{proof}
